Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization

### 2023-10-19 manuscript.Rmd
add declaration for submission
take same affiliation as in software evaluation

---
title: "Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization"
bibliography: references.bib
abstract: "`r paste0(c('\\\\noindent', readLines('abstract.md')), collapse = ' ')`"
shorttitle: "The Objective of Preregistration"
author: 
  - name: Aaron Peikert
    affiliation: "1,2,3"
    email: peikert@mpib-berlin.mpg.de
    corresponding : yes    # Define only one corresponding author
    address: "Center for Lifespan Psychology, Max Planck Institute for Human Development, Lentzeallee 94, 14195 Berlin, Germany"
    role:
      - Conceptualization
      - Writing---Original Draft Preparation
      - Writing---Review & Editing
      - Methodology
      - Formal analysis
      - Software
      - Visualization
      - Project administration
  - name: Maximilian S. Ernst
    affiliation: "1"
    role:
      - Writing---Review & Editing
      - Formal analysis
      - Validation
  - name: Andreas M. Brandmaier
    affiliation: "1, 2, 4"
    role:
      - Writing---Review & Editing
      - Supervisions
affiliation:
  - id: 1
    institution: Center for Lifespan Psychology, Max Planck Institute for Human Development
  - id: 2
    institution: Max Planck UCL Centre for Computational Psychiatry and Ageing Research
  - id: 3
    institution: Department of Psychology, Humboldt-Universität zu Berlin
  - id: 4
    institution: Department of Psychology, MSB Medical School Berlin
keywords: |
  preregistration; confirmation; exploration; hypothesis testing; Bayesian; Open Science
wordcount         : "`r tryCatch(wordcountaddin::word_count(), error = function(x)'~7000')`"
figsintext: false
figurelist: false
tablelist : false
footnotelist: false
lineno: true
linkcolor: "blue"
csquotes: false
class: "man"
documentclass: "apa7"
csl: "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
output:
  papaja::apa6_pdf:
    latex_engine: xelatex
header-includes: |
  \raggedbottom
  \makeatletter

@@

<!--introduce confirmatory vs. exploratory-->
The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis generation and hypothesis testing, or prediction and postdiction  [@hoyningen-hueneContextDiscoveryContext2006; @shmueliExplainPredict2010; @nosekPreregistrationRevolution2018; @tukeyWeNeedBoth1980].
Despite the different names, it is fundamentally the same dichotomy that is at stake here.
There is a broad consensus that both approaches are necessary for science to progress; exploration, to make new discoveries and confirmation, to expose these discoveries to potential falsification, and assess empirical support for the theory.
However, mistaking exploratory findings for empirically confirmed results is dangerous.
It inflates the likelihood of believing that there is evidence supporting a given hypothesis, even if it is false.
A variety of problems, such as researchers' degrees of freedom together with researchers' hindsight bias or naive p-hacking have led to such mistakes becoming commonplace yet unnoticed for a long time.
Recognizing them has led to a crisis of confidence in the empirical sciences [@ioannidisWhyMostPublished2005], and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
As a response to the crisis, evermore researchers preregister their hypotheses and their data collection and analysis plans in advance of their studies [@nosekPreregistrationRevolution2018].
They do so to stress the predictive nature of their registered statistical analyses, often with the hopes of obtaining a label that marks the study as "confirmatory".
Indeed, rigorous application of preregistration prevents researchers from reporting a set of results produced by an arduous process of trial and error as a simple confirmatory story [@wagenmakersAgendaPurelyConfirmatory2012] while keeping low false-positive rates.
This promise of a clear distinction between confirmation and exploration has obvious appeal to many who have already accepted the practice.
Still, the majority of empirical researchers do not routinely preregister their studies.
One reason may be that some do not find that the theoretical advantages outweigh the practical hurdles, such as specifying every aspect of a theory and the corresponding analysis in advance.
We believe that we can reach a greater acceptance of preregistration by explicating a more general objective of preregistration that benefits all kinds of studies, even those that allow data-dependent decisions.

<!--the current goal of preregistration is confirmation-->
One goal of preregistration that has received widespread attention is to clearly distinguish confirmatory from exploratory research [@mellorEasyPreregistrationWill2018; @nosekPreregistrationRevolution2018; @wagenmakersAgendaPurelyConfirmatory2012; @simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
In such a narrative, preregistration is justified by a confirmatory research agenda.
However, two problems become apparent under closer inspection.
First, many researchers do not subscribe to a purely confirmatory research agenda [@tukeyExploratoryDataAnalysis1972; @baumeisterChartingFutureSocial2016; @finkelReplicabilityOtherFeatures2017].
Second, there is no strict mapping of the categories preregistered vs. non-preregistered onto the categories confirmatory vs. exploratory research.

@@

This is the result of researchers applying one of two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration, to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterward.
The latter is a frequent occurrence and, perhaps more worryingly, often remains undisclosed [@akkerEffectivenessPreregistrationPsychology2023; @claesenComparingDreamReality2021].
Both strategies may be used for sensible scientific reasons or with the self-serving intent of generating desirable results.
Thus, insisting on equating preregistration and confirmation has led to the criticism that, all things considered, preregistration is actually harmful and neither sufficient nor necessary for doing good science [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021].

@@

<!---outline-structure--->
To form such an objective for preregistration, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978, but formalized as the probability of proving a hypothesis wrong if it does not hold].

We are interested in why preregistrations should change researchers' evaluation of evidence.
Applying a Bayesian framework allows us to investigate our research question most straightforwardly. 
Specifically, it allows us to model changes in subjective degrees of belief due to preregistration or, more simply, "persuasion".
Please note that our decision to adopt a Bayesian philosophy of science does not make assumptions about the statistical methods researchers use.
In fact, this conceptualization is intentionally as minimal as possible to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.
However, we should note that Popperians would be appalled that we are content with positive inductive inferences (but we dislike "failing to disprove" too much to give it up), and Neopopperians would flinch that we assign probabilities to beliefs (we are fond of calculating things).
While the latter move is not strictly necessary it allows us to connect the more abstract considerations more closely with the behavior of researchers.

We outline two interpretations of preregistration.
The first one corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or, equivalently, by limiting type-I error (when dichotomous decisions about theories are an inferential goal).
We argue that this view on the utility of preregistration can be interpreted as maximizing theoretical risk, which otherwise may be reduced by researchers' degrees of freedom, p-hacking, and suchlike.
The second interpretation is our main contribution:
We argue that contrary to the classic view, the objective of preregistration is *not* the maximization of theoretical risk but rather the minimization of uncertainty about the theoretical risk.
This interpretation leads to a broad applicability of preregistration to both exploratory and confirmatory studies.

@@

The assumption of rationality can be connected to Bayesian reasoning and leads to our adoption of the framework.
Our rationale is as follows.
Researchers who decide to conduct a certain study are actually choosing a study to bet on.
They have to "place the bet" by conducting the study by investing resources and stand to gain epistemic value with some probability.
This conceptualization of choosing a study as a betting problem allows us to apply a "Dutch book" argument [@christensenCleverBookiesCoherent1991].
This argument states that any better must follow the axioms of probability to avoid being "irrational," i.e., accepting bets that lead to sure losses.
Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modeling what liberties researchers have in what studies to conduct.
Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
The axioms of probability are sufficient to derive the Bayes formula, on which we will heavily rely for our further arguments.
The argument is not sufficient, however, to warrant conceptualizing the kind of epistemic value we reason about in terms of posterior probability; that remains a leap of faith.
However, the argument applies to any reward function that satisfies the "statistical relevancy condition" [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974] or as recent psychological literature come to call it "diagnosticity" [@fiedlerWhatConstitutesStrong2017; @oberauerAddressingTheoryCrisis2019].
That is, evidence only increases epistemic value for a theory if the evidence is more likely to be observed under the theory than under the alternative.

@@

The posterior probability $P(H|E)$ is of great relevance since it is often used directly or indirectly as a measure of confirmation of a hypothesis.
In the tradition of Carnap, in its direct use, it is called confirmation as firmness; in its relation to the a priori probability $P(H)$, it is called _increase in firmness_ [@carnapLogicalFoundationsProbability1950, preface to the 1962 edition].
We concentrate on the posterior probability because of its simplicity but take it only as one example of a possible measure.
In reality, surely researchers differ in what function they apply to judge evidence and it is often most fruitful to compare more than two competing hypotheses.
The goal is therefore to reason about the space of possible measures researchers might apply.
However, since it is reasonable that any measure of confirmation increases monotonically with an increase in posterior probability $P(H|E)$, we might well take it to illustrate our reasoning.

In short, we want to increase posterior probability $P(H|E)$.
Increases in posterior probability $P(H|E)$ are associated with increased epistemic value, of which we want to maximize the expectation.
So how can we increase posterior probability?
The Bayes formula yields three components that influence confirmation, namely $P(H)$, $P(E|H)$ and $P(E)$.
The first option leads us to the unsurprising conclusion that higher a priori probability $P(H)$ leads to higher posterior probability $P(H|E)$.
If a hypothesis is more probable to begin with, observing evidence in its favor will result in a hypothesis that is more strongly confirmed, all else being equal.
However, the prior probability of a hypothesis is nothing our study design can change.
The second option is equally reasonable; that is, an increase in $P(E|H)$ leads to a higher posterior probability $P(H|E)$.
$P(E|H)$ is the probability of obtaining evidence for a hypothesis when it holds.
We call this probability of detecting evidence, given that the hypothesis holds "detectability."
Consequently, researchers should ensure that their study design allows them to find evidence for their hypothesis, in case it is true.
When applied strictly within the bounds of null hypothesis testing, detectability is equivalent to power (or the complement of type-II error rate).
However, while detectability is of great importance for study design, it is not directly relevant to what a preregistration is comunicating to other researchers.
However, we latter discuss how issues detectability must be considered for the issue of preregistration.
Thus, $P(E)$ remains to be considered.
Since $P(E)$ is the denominator, decreasing it can increase the posterior probability.
In other words, high risk, high reward.

@@

Let us note some interesting properties of theoretical risk $P(\neg{}E|\neg{}H)$.
First, increasing theoretical risk leads to higher posterior probability $P(H|E)$, our objective.
Second, if the theoretical risk is smaller than detectability $P(E|H)$ it follows that the posterior probability must decrease when observing the evidence.
If detectability exceeds theoretical risk, the evidence is less likely under the theory than it is when the theory does not hold (an inversion of statistical relevancy).
Third, if the theoretical risk equals zero, then posterior probability is at best equal to prior probability but only if detectability is perfect ($P(H|E)$ = 1).
In other words, observing a sure fact does not lend credence to a hypothesis.

The last statement sounds like a truism but is directly related to Popper's seminal criterion of demarcation.
He stated that if it is impossible to prove that a hypothesis is false ($P(\neg{}E|\neg{}H) = 0$, theoretical risk is zero), it cannot be considered a scientific hypothesis [@popperLogicScientificDiscovery2002, p. 18].
We note these relations to underline that the Bayesian rationale we apply here is able to reconstruct many commonly held views on riskiness and epistemic value (but Popper would disagree with attaching any form of believe in $H$ on the basis of evidence beyond failing to disprove it).

@@

Second, available resources limit theoretical risk.
Increasing theoretical risk $P(\neg{}E|\neg{}H)$ will usually decrease detectability $P(E|H)$ unless more resources are invested.
This is similar to the well known tradeoff between type-I error rate and power.
Tasking preregistration with an increase in theoretical risk makes it difficult to balance this trade-off.
Mindlessly maximizing theoretical risk would either never produce evidence or require huge amounts of resources.
As noted before, we want high detectability and high theoretical risk.
Maximizing one at the expense of the other is not necessarily beneficial for increasing epistemic value but depends on the specific function they apply to judge evidence and their specific location on the curve.
One advantage of our framework is that researchers can employ it to balance the trade-off more effectively assuming they are willing to make some simplifiying assumptions.

@@

One widely known factor is the contribution of statistical methods to theoretical risk.
Theoretical risk $P(\neg{}E|\neg{}H)$ is deeply connected with statistical methods, because it is related to the type-I error rate in statistical hypothesis testing $P(E|\neg{}H)$ by $P(\neg{}E|\neg{}H) = 1 - P(E|\neg{}H)$, if you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative-hypothesis because then the nill-hypothesis is $\neg{}H$.
Because many researchers are familiar with the type-I error rate, it can be helpful to remember this connection to theoretical risk.
Researchers who choose a smaller type-I error rate can be more sure of their results, if significant, because the theoretical risk is higher.
However, this connection should not be overinterpreted for two reasons.
First, according to most interpretations of null hypothesis testing, the absence of a significant result should not generally be interpreted as evidence against the hypothesis [@mayoStatisticalInferenceSevere2018, 5.3].
Second, the research hypothesis seldomly equals the statistical alternative-hypothesis (most research hypothesis are more specific then "any value except zero").
In fact, it is entirely possible to assume the null hypothesis as a research hypothesis, as is commonly done in e.g., structural equation modelling, where the roles of detectability, theoretical risk and type-I/II error rate switch.
We argue that theoretical risk (and hence its complement, $P(E|\neg{}H)$) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
Type-I error rate is the property of a statistical test under some assumptions, whereas theoretical risk is a researchers believe.
One may take such theoretical properties as a first starting point to form a substantive believe but surely researchers ought to take more factors into consideration.
For example, if a researcher believes that there might be confounding variables at play for the relation between two variables, that should decrease the theorethical risk; after all they might find an association purely on account of the confounders.

@@

## Sources of uncertainty

@@

Of course, the assigned probabilities and the distribution $Q$ vary from study to study and researcher to researcher (and even the measure of confirmation), but we can illustrate the effect of uncertainty with an example.
Assuming $P(E|H) = 0.8$ (relective of the typically strived for power of 80%).
Let us further assume that the tested hypothesis is considered unlikely to be true by the research community before the study is conducted ($P(H) = 0.1$) and assign a uniform distribution for $P(E|\neg{}H) \sim U([1-\tau, 1])$ where $\tau{}$ is set to $1 - \alpha$, reflecting our assumption that this term gives an upper bound for theoretical risk $P(\neg{}E|\neg{}H)$.
We chose this uniform distribution as it is the maximum entropy distribution with support $[1-\tau, 1]$ and hence conforms to our Bayesian framework [@giffinUpdatingProbabilitiesData2007].

@@

```{r other-measures-fig, cache=TRUE, warning=FALSE}
#| fig.cap: "\\label{fig:other-measures}Several measures for confirmation as an increase in firmness as a function of $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line). Measure taken from @sprengerBayesianPhilosophyScience2019, Table 1.3, p. 51."

@@

The argument for a confirmatory research agenda is that by increasing theoretical risk we increase expected epistemic value, i.e., moving to the right on the x-axis in Figure \ref{fig:posterior} increases posterior probability (on the y-axis).
However, if a hypothesis in a certain study has low theoretical risk, there is not much researchers can do about it.
However, studies do not only differ by how high the theoretical risk is but also by how certain the recipient is about the theoretical risk.
A study that has a very high theoretical risk (e.g., `r (1-confirmatory_tau)*100`% chance that if the hypothesis is wrong, evidence in its favor will be observed,) but has also maximum uncertainty will result in a posterior probability of `r round(confirmatory_uncertain*100)`%, while the same study with maximum certainty will result in `r round(confirmatory_certain*100)`% posterior probability.
The other factors (detectability, prior beliefs, measure of epistemic value) and, therefore, the extent of the benefit varies, of course, with the specifics of the study.
Crucially, even studies with some exploratory aspects benefit from preregistration, e.g., in this scenario with a $\tau$ = `r some_exploration_tau` (false positive rate of `r 1-some_exploration_tau`) moving from uncertain to certain increases the posterior from `r round(some_exploration_uncertain, 2)` to `r round(some_exploration_certain, 2)`.
We find it helpful to calculate some example because of the nonlinear nature of the evidence functions.

@@

Communicating clearly how authors of a scientific report collected their data and consequently analyzed it to arrive at the evidence they present is crucial for judging the theoretical risk they took.
Preregistrations are ideal for communicating just that because any description after the fact is prone to be incomplete.
For instance, the authors could have opted for selective reporting, that is, they decided to exclude a number of analytic strategies they tried out.
That is not to say that every study that was not-preregistered was subjected to practices of questionable research practices.
The point is that we cannot exclude it with certainty.
This uncertainty is drastically reduced if the researchers have described what they intended to do beforehand and then report that they did exactly that.
In that case, readers can be certain they received a complete account of the situation.
They still might be uncertain about the actual theoretical risk the authors took, but to a much smaller extent than if the study would not have been preregistered.

The remaining sources of uncertainty might be unfamiliarity with statistical methods or experimental paradigms used, the probability of an implementation error in the statistical analyses, a bug in the software used for analyses, etc.
To further reduce the uncertainty about theoretical risk, researchers must therefore publish code and ideally data.
After all, computational reproducibility is only possible if the data analytic procedure was communicated clearly enough to allow others to retrace the computational steps [@peikertReproducibleDataAnalysis2021].

In any case, a well-written preregistration should aim to reduce the uncertainty about the theoretical risk and hence increase the persuasiveness of evidence.
Therefore, a study that perfectly adhered to its preregistration will resemble the solid line in Figure \ref{fig:posterior}/\ref{fig:other-measures}.
Crucially, perfect means here that the theoretical risk can be judged with low uncertainty, not that the theoretical risk is necessarily high.

# Hacking, harking, and other harms

The importance of distinguishing between low and highly uncertain theoretical risk becomes perhaps clearer if we consider couple of hypothetical cases.

1. We know with absolute certainty that the results are hacked.
2. The hypothesis was picked to explain the results.
3. We can not exclude the possibility of p-hacking.
4. The results are obtained by planned exploration.
5. The results are obtained by unplanned exploration.

>>>>

---
title: "Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization"
bibliography: references.bib
abstract: "`r paste0(c('\\\\noindent', readLines('abstract.md')), collapse = ' ')`"
shorttitle: "The Objective of Preregistration"
author: 
  - name: Aaron Peikert
    affiliation: "1,2,3"
    email: peikert@mpib-berlin.mpg.de
    corresponding : yes    # Define only one corresponding author
    address: "Center for Lifespan Psychology, Max Planck Institute for Human Development, Lentzeallee 94, 14195 Berlin, Germany"
    role:
      - Conceptualization
      - Writing---Original Draft Preparation
      - Writing---Review & Editing
      - Methodology
      - Formal analysis
      - Software
      - Visualization
      - Project administration
  - name: Maximilian S. Ernst
    affiliation: "1, 4"
    role:
      - Writing---Review & Editing
      - Formal analysis
      - Validation
  - name: Andreas M. Brandmaier
    affiliation: "1, 3, 5"
    role:
      - Writing---Review & Editing
      - Supervisions
affiliation:
  - id            : "1"
    institution   : "Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany"
  - id            : "2"
    institution   : "Department of Imaging Neuroscience, University College London, London, UK"
  - id            : "3"
    institution   : "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany"
  - id            : "4"
    institution   : "Max Planck School of Cognition, Leipzig, Germany"
  - id            : "5"
    institution   : "Department of Psychology, MSB Medical School Berlin, Berlin, Germany"
keywords: |
  preregistration; confirmation; exploration; hypothesis testing; Bayesian; Open Science
wordcount         : "`r tryCatch(wordcountaddin::word_count(), error = function(x)'~7000')`"
figsintext: false
figurelist: false
tablelist : false
footnotelist: false
lineno: true
linkcolor: "blue"
csquotes: false
class: "man"
documentclass: "apa7"
csl: "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
output:
  papaja::apa6_pdf:
    latex_engine: xelatex
header-includes: |
  \raggedbottom
  \makeatletter

@@

<!--introduce confirmatory vs. exploratory-->
The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis generation and hypothesis testing, or prediction and postdiction  [@hoyningen-hueneContextDiscoveryContext2006; @shmueliExplainPredict2010; @nosekPreregistrationRevolution2018].
Despite the different names, it is fundamentally the same dichotomy that is at stake here.
There is a broad consensus that both approaches are necessary for science to progress; exploration, to make new discoveries and confirmation, to expose these discoveries to potential falsification, and assess empirical support for the theory.
However, mistaking exploratory findings for empirically confirmed results is dangerous.
It inflates the likelihood of believing that there is evidence supporting a given hypothesis, even if it is false.
A variety of problems, such as researchers' degrees of freedom together with researchers' hindsight bias or naive p-hacking have led to such mistakes becoming commonplace yet unnoticed for a long time.
Recognizing them has led to a crisis of confidence in the empirical sciences [@ioannidisWhyMostPublished2005], and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
As a response to the crisis, evermore researchers preregister their hypotheses and their data collection and analysis plans in advance of their studies [@nosekPreregistrationRevolution2018].
They do so to stress the predictive nature of their registered statistical analyses, often with the hopes of obtaining a label that marks the study as "confirmatory".
Indeed, rigorous application of preregistration prevents researchers from reporting a set of results produced by an arduous process of trial and error as a simple confirmatory story [@wagenmakersAgendaPurelyConfirmatory2012] while keeping low false-positive rates.
This promise of a clear distinction between confirmation and exploration has obvious appeal to many who have already accepted the practice.
Still, the majority of empirical researchers do not routinely preregister their studies.
One reason may be that some do not find that the theoretical advantages outweigh the practical hurdles, such as specifying every aspect of a theory and the corresponding analysis in advance.
We believe that we can reach a greater acceptance of preregistration by explicating a more general objective of preregistration that benefits all kinds of studies, even those that allow data-dependent decisions.

<!--the current goal of preregistration is confirmation-->
One goal of preregistration that has received widespread attention is to clearly distinguish confirmatory from exploratory research [@mellorEasyPreregistrationWill2018; @nosekPreregistrationRevolution2018; @wagenmakersAgendaPurelyConfirmatory2012; @simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
In such a narrative, preregistration is justified by a confirmatory research agenda.
However, two problems become apparent under closer inspection.
First, many researchers do not subscribe to a purely confirmatory research agenda.
Second, there is no strict mapping of the categories preregistered vs. non-preregistered onto the categories confirmatory vs. exploratory research.

@@

This is the result of researchers applying one of two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration, to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterward.
Both strategies may be used for sensible scientific reasons or with the self-serving intent of generating desirable results.
Thus, insisting on equating preregistration and confirmation has led to the criticism that, all things considered, preregistration is actually harmful and neither sufficient nor necessary for doing good science [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021].

@@

<!---outline-structure--->
To form such an objective for preregistration, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978, but formalized as the probability of proving a hypothesis wrong if it does not hold], which is inversely related to the type-I error rate in null hypothesis testing.

Further, we outline two interpretations of preregistration.
The first one corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or, equivalently, by limiting type-I error (when dichotomous decisions about theories are an inferential goal).
We argue that this view on the utility of preregistration can be interpreted as maximizing theoretical risk, which otherwise may be reduced by researchers' degrees of freedom, p-hacking, and suchlike.
The second interpretation is our main contribution:
We argue that contrary to the classic view, the objective of preregistration is *not* the maximization of theoretical risk but rather the minimization of uncertainty about the theoretical risk.
This interpretation leads to a broad applicability of preregistration to both exploratory and confirmatory studies.

@@

The assumption of rationality can be connected to Bayesian reasoning and leads to our adoption of the framework.
Our rationale is as follows.
Researchers who decide to conduct a certain study are actually choosing a study to bet on.
They have to "place the bet" by conducting the study by investing resources and stand to gain epistemic value with some probability.
This conceptualization of choosing a study as a betting problem allows us to apply a "Dutch book" argument [@christensenCleverBookiesCoherent1991].
This argument states that any better must follow the axioms of probability to avoid being "irrational," i.e., accepting bets that lead to sure losses.
Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modeling what liberties researchers have in what studies to conduct.
Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
The axioms of probability are sufficient to derive the Bayes formula, on which we will heavily rely for our further arguments.
The argument is not sufficient, however, to warrant conceptualizing the kind of epistemic value we reason about in terms of posterior probability; that remains a leap of faith.
However, the argument applies to any reward function that satisfies the "statistical relevancy condition" [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974].
That is, evidence only increases epistemic value for a theory if the evidence is more likely to be observed under the theory than under the alternative.

Please note that our decision to adopt this aspect of the Bayesian philosophy of science does not make assumptions about the statistical methods researchers use.
In fact, this conceptualization is intentionally as minimal as possible to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.

@@

The posterior probability $P(H|E)$ is of great relevance since it is often used directly or indirectly as a measure of confirmation of a hypothesis.
In the tradition of Carnap, in its direct use, it is called confirmation as firmness; in its relation to the a priori probability $P(H)$, it is called _increase in firmness_ @carnapLogicalFoundationsProbability1950, preface to the 1962 edition].
As noted before, we concentrate on posterior probability as a measure of epistemic value since no measure shows universally better properties than others.
However, it is reasonable that any measure of confirmation increases monotonically with an increase in posterior probability $P(H|E)$, and our argument applies to those measures as well.

In short, we want to increase posterior probability $P(H|E)$.
Increases in posterior probability $P(H|E)$ are associated with increased epistemic value, of which we want to maximize the expectation.
So how can we increase posterior probability?
The Bayes formula yields three components that influence confirmation, namely $P(H)$, $P(E|H)$ and $P(E)$.
The first option leads us to the unsurprising conclusion that higher a priori probability $P(H)$ leads to higher posterior probability $P(H|E)$.
If a hypothesis is more probable to begin with, observing evidence in its favor will result in a hypothesis that is more strongly confirmed, all else being equal.
However, the prior probability of a hypothesis is nothing our study design can change.
The second option is equally reasonable; that is, an increase in $P(E|H)$ leads to a higher posterior probability $P(H|E)$.
$P(E|H)$ is the probability of obtaining evidence for a hypothesis when it holds.
We call this probability of detecting evidence, given that the hypothesis holds "detectability."
Consequently, researchers should ensure that their study design allows them to find evidence for their hypothesis, in case it is true.
When applied strictly within the bounds of null hypothesis testing, detectability is equivalent to power (or the complement of type-II error rate).
However, while detectability is of great importance for study design, it is not directly relevant to the objective of preregistration.
Thus, $P(E)$ remains to be considered.
Since $P(E)$ is the denominator, decreasing it can increase the posterior probability.
In other words, high risk, high reward.

@@

Let us note some interesting properties of theoretical risk $P(\neg{}E|\neg{}H)$.
First, increasing theoretical risk leads to higher posterior probability $P(H|E)$, our objective.
Second, if the theoretical risk is smaller than detectability $P(E|H)$ it follows that the posterior probability must decrease when observing the evidence.
If detectability exceeds theoretical risk, the evidence is less likely under the theory than it is when the theory does not hold.
Third, if the theoretical risk equals zero, then posterior probability is at best equal to prior probability but only if detectability is perfect ($P(H|E)$ = 1).
In other words, observing a sure fact does not lend credence to a hypothesis.

The last statement sounds like a truism but is directly related to Popper's seminal criterion of demarcation.
He stated that if it is impossible to prove that a hypothesis is false ($P(\neg{}E|\neg{}H) = 0$, theoretical risk is zero), it cannot be considered a scientific hypothesis [@popperLogicScientificDiscovery2002, p. 18].
We note these relations to underline that the Bayesian rationale we apply here is able to reconstruct many commonly held views on riskiness and epistemic value.

@@

Second, available resources limit theoretical risk.
Increasing theoretical risk $P(\neg{}E|\neg{}H)$ will usually decrease detectability $P(E|H)$ unless more resources are invested.
In other words, one cannot increase power while maintaining the same type-I error rate without increasing the invested resources.
Tasking preregistration with an increase in theoretical risk makes it difficult to balance this trade-off.
Mindlessly maximizing theoretical risk would either never produce evidence or require huge amounts of resources.

@@

One widely known factor is the contribution of statistical methods to theoretical risk.
Theoretical risk $P(\neg{}E|\neg{}H)$ is deeply connected with statistical methods, because it is related to the type-I error rate in statistical hypothesis testing $P(E|\neg{}H)$ by $P(\neg{}E|\neg{}H) = 1 - P(E|\neg{}H)$, if you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative-hypothesis because then the nill-hypothesis is $\neg{}H$.
Because many researchers are familiar with the type-I error rate, it can be helpful to remember this connection to theoretical risk.
Researchers who choose a smaller type-I error rate can be more sure of their results, if significant, because the theoretical risk is higher.
However, this connection should not be overinterpreted for two reasons.
First, according to most interpretations of null hypothesis testing, the absence of a significant result should not generally be interpreted as evidence against the hypothesis [@mayoStatisticalInferenceSevere2018, 5.3].
Second, the research hypothesis seldomly equals the statistical alternative-hypothesis.
We argue that theoretical risk (and hence its complement, $P(E|\neg{}H)$) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.

@@

## Sources of Uncertainty

@@

Of course, the assigned probabilities and the distribution $Q$ vary from study to study and researcher to researcher, but we can illustrate the effect of uncertainty with an example.
Assuming $P(E|H) = 0.8$ (relective of the typically strived for power of 80%).
Let us further assume that the tested hypothesis is considered unlikely to be true by the research community before the study is conducted ($P(H) = 0.1$) and assign a uniform distribution for $P(E|\neg{}H) \sim U([1-\tau, 1])$ where $\tau{}$ is set to $1 - \alpha$, reflecting our assumption that this term gives an upper bound for theoretical risk $P(\neg{}E|\neg{}H)$.
We chose this uniform distribution as it is the maximum entropy distribution with support $[1-\tau, 1]$ and hence conforms to our Bayesian framework [@giffinUpdatingProbabilitiesData2007].

@@

```{r other-measures-fig, cache=TRUE, warning=FALSE}
#| fig.cap: "\\label{fig:other-measures}Several measures for confirmation as an increase in firmness as a function of $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line)."

@@

The argument for a confirmatory research agenda is that by increasing theoretical risk we increase expected epistemic value, i.e., moving to the right on the x-axis in Figure \ref{fig:posterior} increases posterior probability (on the y-axis).
However, if a hypothesis in a certain study has low theoretical risk, there is not much researchers can do about it.
However, studies do not only differ by how high the theoretical risk is but also by how certain the recipient is about the theoretical risk.
A study that has a very high theoretical risk (e.g., `r (1-confirmatory_tau)*100`% chance that if the hypothesis is wrong, evidence in its favor will be observed,) but has also maximum uncertainty will result in a posterior probability of `r round(confirmatory_uncertain*100)`%, while the same study with maximum certainty will result in `r round(confirmatory_certain*100)`% posterior probability.
The other factors (detectability, prior beliefs, measure of epistemic value) and, therefore, the extent of the benefit varies, of course, with the specifics of the study.
Crucially, even studies with some exploratory aspects benefit from preregistration, e.g., in this scenario with a $\tau$ = `r some_exploration_tau` (false positive rate of `r 1-some_exploration_tau`) moving from uncertain to certain increases the posterior from `r round(some_exploration_uncertain, 2)` to `r round(some_exploration_certain, 2)`.

@@

Communicating clearly how authors of a scientific report collected their data and consequently analyzed it to arrive at the evidence they present is crucial for judging the theoretical risk they took.
Preregistrations are ideal for communicating just that because any description after the fact is prone to be incomplete.
For instance, the authors could have opted for selective reporting, that is, they decided to exclude a number of analytic strategies they tried out.
That is not to say that every study that was not-preregistered was subjected to practices of questionable research practices.
The point is that we cannot exclude it with certainty.
This uncertainty is drastically reduced if the researchers have described what they intended to do beforehand and then report that they did exactly that.
In that case, readers can be certain they received a complete account of the situation.
They still might be uncertain about the actual theoretical risk the authors took, but to a much smaller extent than if the study would not have been preregistered.
The remaining sources of uncertainty might be unfamiliarity with statistical methods or experimental paradigms used, the probability of an implementation error in the statistical analyses, a bug in the software used for analyses, etc.
In any case, a well-written preregistration should aim to reduce the uncertainty about the theoretical risk and hence increase the persuasiveness of evidence.
Therefore, a study that perfectly adhered to its preregistration will resemble the solid line in Figure \ref{fig:posterior}/\ref{fig:other-measures}.
Crucially, perfect means here that the theoretical risk can be judged with low uncertainty, not that the theoretical risk is necessarily high.
